---
title: "Task 3 - Text Analysis"
author: "Deanna Elliott"
date: "3/13/2022"
output: 
  html_document:
    theme: spacelab
    code_folding: hide
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(tidytext)
library(textdata)
library(pdftools)
```

### Overview

This report explores the word count and sentimentality of the song *The Greatest Show on Earth* by Finnish symphonic metal Nightwish. The song is from their eighth studio album *Endless Forms Most Beautiful*, released in 2015, and chronicles the evolution of Earth. Text wrangling is used to find counts of the top 5 words in the song, and two sentiment lexicons are used to describe the overall tone of each movement of the song.

**Text Source:** LyricFind.com

<br>

### Text Wrangling

```{r}

# read in text

gsoe_text <- pdf_text(here("data", "gsoe_nightwish.pdf"))

# break down into pages and lines

gsoe_lines <- data.frame(gsoe_text) %>% 
  mutate(page = 1:n()) %>% 
  mutate(full_text = str_split(gsoe_text, pattern = '\\n')) %>% 
  unnest(full_text) %>% 
  mutate(full_text = str_squish(full_text))

```

```{r}

# tidy: separate the movements

gsoe_mvmts <- gsoe_lines %>% 
  slice(-(0:8)) %>% 
  mutate(movement = ifelse(str_detect(full_text, 'Movement'), full_text, NA)) %>%
  fill(movement, .direction = 'down') %>% 
  separate(col = movement, into = c('mvmt', 'no'), sep = " ")

# the code won't recognize Movement 1; I tried multiple times, so instead...

gsoe_mvmts$no[is.na(gsoe_mvmts$no)] <- 1

```

```{r}

# word counts

gsoe_words <- gsoe_mvmts %>% 
  unnest_tokens(word, full_text, token = 'words') %>% 
  select(-gsoe_text)

gsoe_wordcount <- gsoe_words %>% 
  count(mvmt, no, word)

# get rid of stop words

gsoe_words_clean <- gsoe_words %>% 
  anti_join(stop_words, by = 'word')

nonstop_count <- gsoe_words_clean %>% 
  count(mvmt, no, word)

```

```{r}

# find top 5 words in whole song

top_5_words <- nonstop_count %>% 
  group_by(no) %>% 
  arrange(-n) %>% 
  slice(1:5) %>% 
  ungroup()

ggplot(data = top_5_words) +
  geom_col(aes(x = n, y = reorder(word, -n)), fill = 'royalblue') +
  labs(x = "Count", y = "Word") +
  theme_minimal()
  
```

**Figure 1. Top words.** This graph ostensibly shows the top five words in the song *The Greatest Show on Earth* by Nightwish. Many words were repeated an equal number of times, so there are a total of 19 words that tie for top five.

<br>

### Sentiment Analysis

```{r}

afinn <- get_sentiments(lexicon = 'afinn')

gsoe_afinn <- gsoe_words_clean %>% 
  inner_join(get_sentiments('afinn'), by = 'word')
```

```{r}

afinn_counts <- gsoe_afinn %>% 
  count(no, value)

afinn_means <- gsoe_afinn %>% 
  group_by(no) %>% 
  summarize(mean_afinn = mean(value))

ggplot(data = afinn_means, 
       aes(x = no, y = mean_afinn)) +
  geom_col(fill = 'royalblue', color = 'black') +
  labs(x = "Movement", y = "Mean Afinn Value") +
  theme_minimal()
```

**Figure 2. Afinn values by movement.** This graph shows the mean value of words in each of the song's four movements based on the Afinn sentiment lexicon. Higher values are associated with positive words. Movement 3 had a mean value of zero.

<br>

```{r}

nrc <- get_sentiments(lexicon = 'nrc')

gsoe_nrc <- gsoe_words_clean %>% 
  inner_join(get_sentiments('nrc'))

gsoe_nrc$no[is.na(gsoe_nrc$no)] <- 1


gsoe_nrc_counts <- gsoe_nrc %>% 
  count(no, sentiment)

ggplot(data = gsoe_nrc_counts,
       aes(x = no, y = n)) +
  geom_col(fill = 'royalblue', color = 'black') +
  facet_wrap(~sentiment) +
  theme_minimal() +
  labs(x = 'Movement', y = 'Word Count')
```


**Figure 3. Sentiments by movement.** The above graph shows the count of words per movement which are associated with 10 different emotions and/or sentiments. The analysis is based on the NRC sentiment lexicon.

<br>
