---
title: "Task 1 - Spatial Data Visualization"
author: "Deanna Elliott"
date: "3/1/2022"
output: 
  html_document:
    theme: spacelab
    code_folding: hide
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(broom)
library(sf)
library(tmap)
library(gstat)
library(stars)
```

### Overview

This report visualizes sites and counts of oil spills in California in 2008. Data from CA DFW Oil Spill Incident Tracking is used to create an interactive map of each spill site, and a second static map showing the number of spills per county.

<br>

```{r}

# read in spill site file

spill_data_sf <- read_sf(here("data", "spill", "ds394.shp")) %>% 
  janitor::clean_names()

spill_site_sf <- spill_data_sf %>% 
  select(localecity, dateofinci, inlandmari) #sticky geography 

spill_site_df <- spill_site_sf %>% 
  as.data.frame() %>% 
  select(-geometry)

# read in CA counties file

ca_counties_sf <- read_sf(here("data", "spill", 
                               "CA_Counties_TIGER2016.shp"))

ca_subset_sf <- ca_counties_sf %>% 
  janitor::clean_names() %>% 
  select(county_name = name, land_area = aland)
```

```{r}
 # check crs

spill_crs <- spill_site_sf %>% st_crs()
ca_crs <- ca_subset_sf %>% st_crs()

ca_crs_sf <- st_transform(ca_subset_sf, st_crs(spill_site_sf))
```

```{r}
tmap_mode(mode = 'view')

tm_shape(ca_crs_sf) +
  tm_borders(col = 'black') +
  tm_shape(spill_site_sf) +
  tm_dots(col = 'red',
          popup.vars = c("City" = "localecity", "Date" = "dateofinci",
                         "Location" = "inlandmari"))
```

**Figure 1. Oil spills in California.** This interactive map shows individual instances of both marine and inland oil spills throughout California in 2008.

<br>

```{r}

inland_spill_sf <- spill_data_sf %>% 
   janitor::clean_names() %>% 
  mutate(year = lubridate::year(dateofinci)) %>% 
  select(year, inlandmari, localecoun) %>% 
  filter(inlandmari == "Inland")
```

```{r}

# find inland spills per county in 2008

inland_crs_sf <- st_transform(inland_spill_sf, st_crs(ca_subset_sf))

inland_join_sf <- ca_subset_sf %>% 
  st_join(inland_crs_sf)

inland_counts_sf <- inland_join_sf %>% 
  group_by(localecoun) %>% 
  summarize(n_records = sum(!is.na(inlandmari)))
```

```{r}

ggplot(data = inland_counts_sf) +
  geom_sf(aes(fill = n_records), color = 'white', size = 0.1) +
  scale_fill_gradientn(colors = c('lightgrey', 'orange', 'red')) +
  theme_void() +
  labs(fill= 'Number of Oil Spills')
```

**Figure 2. Inland Oil Spills.** This map shows the number of inland oil spills in California, by county, for the year 2008.

<br>

**Data Sources:** California County shapefile (ca_counties_tiger_2016.shp) from the US Census Bureau’s 2016 MAF/TIGER database (accessed [here](https://data.ca.gov/dataset/ca-geographic-boundaries/resource/b0007416-a325-4777-9295-368ea6b710e6?inner_span=True)).

Oil spill data from California Department of Fish & Wildlife and the California State Geoportal - Oil Spill Incident Tracking [ds394] (accessed [here](https://gis.data.ca.gov/datasets/7464e3d6f4924b50ad06e5a553d71086_0/explore?location=36.977441%2C-119.422009%2C6.16&showTable=true)).
